---
title: "Assign. 1 STA 445"
author: "Connor Phillips"
date: "`2/20/2024`"
output: pdf_document
---

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
```

## Directions:
This assignment covers chapter 5.  Please show all work in this document and knit your final draft into a pdf.  This is assignment is about statistical models, which will be helpful if you plan on taking STA 570, STA 371, or STA 571. 

## \color{black} Problem 1: Two Sample t-test

\color{blue} 

a. Load the `iris` dataset.  
```{r}
head(iris)
```

b. Create a subset of the data that just contains rows  for the two species setosa and versicolor using filter.  Use slice_sample to print out 20 random rows of the dataset.
```{r}
iris.subset = iris %>%
  filter(Species == 'setosa' | Species == 'versicolor')
slice_sample(iris.subset, n=20)
```

c. Create a box plot of the petal lengths for these two species using ggplot. Does it look like the mean petal length varies by species?

  Yes, the mean petal length of the setosas (1.5) is significantly smaller than the mean petal length of the versicolors   (4.3).

```{r, fig.width=10}
ggplot(iris.subset, aes(y=Species, x=Petal.Length)) +
  geom_boxplot()

iris.subset %>%
  group_by(Species) %>%
  summarise(mean.Length = mean(Petal.Length))
```

d. Do a two sample t-test using t.test to determine formally if the petal lengths differ. Note:  The book uses the tidy function in the broom package to make the output "nice".  I hate it!  Please don't use tidy.
```{r}
t.test(data=iris.subset, Petal.Length ~ Species, conf.level=0.9)
```

d. What is the p-value for the test? What do you conclude?

  The p-value is 2.2e-16, which is very low and means we can reject the null hypothesis.
 
e. Give a 95% confidence interval for the difference in the mean petal lengths.  
```{r}
t.test(data=iris.subset, Petal.Length ~ Species, conf.level=0.95)
```
f. Give a 99% confidence interval for the difference in mean petal lengths.  (Hint: type ?t.test.  See that you can change the confidence level using the option conf.level)
```{r}
t.test(data=iris.subset, Petal.Length ~ Species, conf.level=0.99)
```
g. What is the mean petal length for setosa?

  The mean petal length of the setosas is 1.5.
 
h. What is the mean petal length for versicolor?
 
  The mean petal length of the versicolors is 4.3.

## Problem 2: ANOVA 

Use the iris data with all three species.

a. Create a box plot of the petal lengths for all three species using ggplot.Does it look like there are differences in the mean petal lengths?
  
  Yes, virginica has a higher mean pedal length than either of the previous two species. 
  
```{r}
ggplot(iris, aes(y=Species, x=Petal.Length)) +
  geom_boxplot() 
```
b. Create a linear model where sepal length is modeled by species. Give it
an appropriate name.
```{r}
iris.model = lm(data=iris, Sepal.Length ~ Species -1)
```

c. Type anova(your model name) in a code chunk.
```{r}
anova(iris.model)
```

d. What is the p-value for the test?  What do you conclude.
  
  The p-value is 2.2e-16, which is very low, and we can reject the null hypothesis.
  
e. Type summary(your model name) in a code chunk.
```{r}
summary(iris.model)
```
f. What is the mean petal length for the species setosa?

  The mean for the species setosa is 5.01.

g. What is the mean petal length for the species versicolor?

  The mean petal length for the species versicolor is 5.94.


## Problem 3: Regression 

Can we describe the relationship between petal length and petal width?

a. Create a scatterplot with petal length on the y-axis and petal width on the x-axis using ggplot.
```{r}
ggplot(iris, aes(x=Petal.Width, y=Petal.Length)) +
  geom_point()
```

b. Create a linear model to model petal length with petal width (length is the response variable and width is the explanatory variable) using lm.
```{r}
petal.model = lm(data=iris, Petal.Length ~ Petal.Width)
petal.model
```

c. What is the estimate of the slope parameter?

  The estimate of the slope parameter is 2.230.

d. What is the estimate of the intercept parameter?

  The estimate of the intercept parameter is 1.084.

e. Use summary() to get additional information.
```{r}
summary(petal.model)
```

## Problem 4: Modeling Trees

Using the `trees` data frame that comes pre-installed in R, follow the steps below to fit the regression model that uses the tree `Height` to explain the `Volume` of wood harvested from the tree.

a. Create a scatterplot of the data using ggplot.
```{r}
ggplot(data=trees, aes(x=Height, y=Volume)) +
  geom_point()
```

b. Fit a `lm` model using the command `model <- lm(Volume ~ Height, data=trees)`.
```{r}
tree.model = lm(data=trees, Volume ~ Height)
tree.model
```

c. Print out the table of coefficients with estimate names, estimated value, standard error, and upper and lower 95% confidence intervals.
```{r}
summary(tree.model)$coef
confint(tree.model, level=0.95)
```

d.Add the model fitted values to the `trees` data frame along with the regression model confidence intervals.  Note: the book does this in a super convoluted way.  Don't follow the model in the book.  Instead try cbind.
```{r}
trees.w.pred = cbind(trees, predict(tree.model, interval = "confidence"))
head(trees.w.pred)
```

e. Graph the data and fitted regression line and uncertainty ribbon.
```{r, fig.width=8, fig.height=8}
ggplot(data = trees.w.pred, aes(x=Height, y=Volume)) +
  geom_point(size=3) +
  geom_line(aes(y=fit)) +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.5, fill = "gold")
```

f. Add the R-squared value as an annotation to the graph using annotate.
```{r, fig.width=8, fig.height=8}
tree.r.squared = summary(tree.model)$r.squared
tree.r.sq = paste('Rsq =', round(tree.r.squared, digits = 3)) 

ggplot(data = trees.w.pred, aes(x=Height, y=Volume)) +
  geom_point(size=3) +
  geom_line(aes(y=fit)) +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.5, fill = "gold") +
  annotate('label', label=tree.r.sq, x=80, y=0, size=5)
```

