---
title: "STA 445 HW2"
author: "Connor Phillips"
date: "2/27/2024"
output: pdf_document
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Problem 1

Create a vector of three elements (2,4,6) and name that vector `vec_a`. Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors together and name the result `vec_c`.
```{r}
vec_a = c(2,4,6)
vec_b = c(8,10,12)
vec_c = vec_a + vec_b
vec_c
```

## Problem 2

Create a vector, named `vec_d`, that contains only two elements (14,20). Add this vector to `vec_a`. What is the result and what do you think R did (look up the recycling rule using Google)? What is the warning message that R gives you?
  The result is 16, 24, 20
  I get the message: "Warning: longer object length is not a multiple of shorter object length"
```{r}
vec_d = c(14,20)
vec_a + vec_d
```

## Problem 3

Next add 5 to the vector vec_a. What is the result and what did R do? Why doesn't in give you a warning message similar to what you saw in the previous problem?

  The result is 7 9 11
  There is no error because 5 is being added to each number in vec_a, it isn't running into the issue of adding numbers within vectors that are two different lengths.
```{r}
vec_a + 5
```

## Problem 4

Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two different ways. 

a.  First using the `seq()` function
```{r}
vec1_5a = seq(1,5,1)
vec1_5a
```

b. Using the `a:b` shortcut.
```{r}
vec1_5b = 1:5
vec1_5b
```


## Problem 5

Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 

a. Using the seq() function  
```{r}
vec2_20a = seq(2,20, by=2)
vec2_20a
```


b. Using the a:b shortcut and some subsequent algebra. 
```{r}
vec2_20b = 2*(1:10)
vec2_20b
```


## Problem 6

Generate a vector of 21 elements that are evenly placed between 0 and 1 using the `seq()` command and name this vector `x`. 
```{r}
x = seq(0, 1, length.out=21)
x
```


## Problem 7

Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ using the `rep()` command to replicate the vector c(2,4,8). 
```{r}
vec2_8rep = rep(c(2, 4, 8), 3)
vec2_8rep
```
  


## Problem 8
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
using the `rep()` command. You might need to check the help file for rep() to see all of the options that rep() will accept. In particular, look at the optional argument `each=`.
```{r}
vec_numRep = rep(c(2, 4, 8), each=4)
vec_numRep
```

## Problem 9 

In this problem, we will work with the matrix 
$$ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]$$
    
a.  Create the matrix in two ways and save the resulting matrix as `M`.

i. Create the matrix using some combination of the `seq()` and `matrix()` commands.
```{r}
M = matrix(seq(2, 30, by=2), nrow=3, ncol=5, byrow = TRUE)
M
```
ii. Create the same matrix by some combination of multiple `seq()` commands and either the `rbind()` or `cbind()` command.
```{r}
c1 = seq(2,10, by=2)
c2 = seq(12,20, by=2)
c3 = seq(22,30, by=2)

M = rbind(c1, c2, c3)
M
```

b. Extract the second row out of M.
```{r}
M[2,]
```

c. Extract the element in the third row and second column of M
```{r}
M[[3, 2]]
```


## Problem 10 

The following code creates a `data.frame` and then has two different methods for removing the rows with `NA` values in the column `Grade`. Explain the difference between the two.

  The first code creates a data frame that is taking the inverse of which grades in the data frame are not true (which indices are false.)
  The second code creates a data frame that is is selecting grades where the index is  not equal to NA. 
```{r, eval=FALSE}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
df[ -which(  is.na(df$Grade) ), ]

df[  which( !is.na(df$Grade) ), ]
```


## Problem 11 

Create and manipulate a list.

a.  Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131
```{r}
x = c(4,5,6,7,8,9,10)
y = c(34,35,41,40,45,47,51)
slope = 2.82
p.value = 0.000131
my.list = list(x=x, y=y, slope=slope, p.value=p.value)
my.list
```

b.  Extract the second element in the list.
```{r}
my.list$y
```


c.  Extract the element named `p.value` from the list.
```{r}
my.list$p.value
```



